---
title: "bernoulli-distribution"
author: "Ajay Karthick"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Coin flip simulation

```{r coin_flip_simulation}
coin <- c('head', 'tail')
num_flips <- 1
sample(coin, size = num_flips, replace = TRUE, prob = c(0.6, 0.4))
```

### Probability Mass Function (PMF):

\[
    f(k; p) = \left\{\begin{array}{lr}
        p, & \text{for } k = 1\\
        q = (1-p), & \text{for } k = 0\\
        \end{array}\right\}
\]

Also expressed as,

\[
  f(k; p) = p^k (1-p)^{(1-k)} \text{ for } k \in \left\{0, 1\right\}
\]

or as,

\[ 
  f(k; p) = pk + (1-p)(1-k) \text{ for } k \in \left\{ 0, 1 \right\}
\]


### Mean / Expected Value

$$\mu = p$$

### Variance
$$\sigma^2 = p (1 - p) = pq$$

#### Calculation of Variance

$E[X^2] = Pr[X = 1] . 1 + Pr[X = 0] . 0 = Pr[X=1] = p$\
$E[X]^2 = p^2$\
$Var[X] = \sigma^2 = E[X^2] - E[X]^2 = p - p^2 = p (1-p) = pq$\

* Binomial Distribution is the distribution of the number of successes in n-independent Bernoulli trials

* Geometric Distribution is the distribution of the the number of trials to get the first success in the n-independent Bernoulli trials.

* Negative Binomial Distribution is the distribution of the number of trials to get the $r^{th}$ success in the n-independent Bernoulli trials.
