---
title: "geometric-distribution"
author: "Ajay Karthick"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Geometric Distribution is the distribution of the number of trials needed to get the first success in repeated independent Bernoulli trials.

For the first success to occur on the $x^{th}$ trial:
1. The first $x-1$ trials must be failures (Probability for this to happen is $(1-p)^{(x-1)}$)
2. $x^{th}$ trial must be a success (Probability for this to happen is $p$)

Hence, 
\[P[X=x] = (1-p)^{(x-1)}p \\
\text{ for x = 1, 2, 3,...}
\]


The Minimum Value the Geometrically distributed random variable X can take is 1.

### Expected Value

$$E[X] = \mu = \frac{1}{p}$$

#### Expected Value Calculation Proof

<iframe  width="560" height="315" src="https://www.youtube.com/embed/7br_-emGNec" data-external= "1" > </iframe>


### Variance

$$Var[X] = \sigma^2 = \frac{1-p}{p^2}$$

#### Variance Calculation Proof

<iframe  width="560" height="315" src="https://www.youtube.com/embed/yTW5uNVh8dw" data-external= "1" > </iframe>


### Probability Mass Function (PMF)

$$P[X=x] = (1-p)^{x-1} * p$$

### Cumulative Density Function (CDF)

\[F(x) = P[X \leq x] = 1 - (1-p)^x\\
\text{for x = 1, 2, 3, ....}
\]

#### CDF Proof

<iframe  width="560" height="315" src="https://www.youtube.com/embed/zq9Oz82iHf0?start=436" data-external= "1" > </iframe>


### rgeom

```{r}
n=10
p=0.4
rgeom(n=n, prob=p)
```

### dgeom

```{r}
x=0:100
p=0.1
pmf <- dgeom(x=x, prob=p)
plot(x, 
     pmf,
     type='h',
     main='Geometric Distribution - PMF',
     xlab='First Success Trial',
     ylab='Probability',
     lwd=3)
```


### pgeom

```{r}
q=0:100
p=0.1
cdf <- pgeom(q=q, prob=p)
plot(q, 
     cdf,
     type='h',
     main='Geometric Distribution - CDF',
     xlab='First Success Trial',
     ylab='Probability',
     lwd=3)
```

### qgeom

```{r}
quantiles <- seq(0, 1, .001)
inverse_cdf <- qgeom(p=quantiles, prob=0.5)
plot(quantiles, 
     inverse_cdf,
     type='s',
     main='Geometric Distribution - Quantile',
     xlab='p',
     ylab='Q(p)',
     lwd=1)
```

