---
title: "Poisson Distribution"
author: "Ajay Karthick"
date: "2022-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<iframe  width="560" height="315" src="https://www.youtube.com/embed/TD1N4hxqMzY" data-external= "1" > </iframe>

Suppose,

* Events are occurring independently.
* The probability that an event occurs in a given length of time does not change through time.

Then X, the number of events in a fixed unit of time, has a Poisson distribution.

### Poisson Paradigm

* Events $A_1, A_2, ...,A_n$, Events are independent or weekly dependent,
* $n$ large,
* $P(A_j)=P_j\text{, }$
* $P_j$'s are small

Then, the claim is that the # of $A_j$'s that occur is approximately Poisson - $Pois(\lambda)$
\[\lambda=\sum_{j=1}^{n}P_j\]

### PMF

\[
P(X = x) = \frac{\lambda^x e^{-\lambda}}{x!} \\
\text{ for x = 0, 1, 2,...}
\]

$e$ - important mathematical constant, base of natural algorithms. It is approximately equal to 2.71828 but it is an irrational number that has infinite non-repeating decimal places.


### Mean

$$\mu = \lambda$$

### Variance

$$\sigma^2 =\lambda$$


### Connection between Binomial and Poisson Distributions

Binomial converges to the Poisson

\[
X \sim Bin(n, p) \text{, let } n \rightarrow \infty \text{, } p \rightarrow 0 \text{, }
\]

\[
\lambda = np \text{ is held constant. So, p is going to 0 at the same rate as n is going to infinity.} 
\]

\[
\text{find what happens to }P(X=k) = {n \choose k} p^k (1 - p)^{n-k} \text{, k is fixed to look at the one specific value for the PMF since } n \rightarrow \infty \text{ and } p \rightarrow 0
\]

\[
\text{Let's write everything in terms of n.} 
\]

\[
\text{Since, } \lambda = np \implies p = \frac{\lambda}{n}.
\]


\[
\text{So, } P[X=k] = \frac{n . (n-1) . (n-2) . .... (n-k+1)}{k!} \frac{\lambda^k}{n^k} \biggl(1 - \frac{\lambda}{n} \biggr)^{n-k} 
\]
\[
= \frac{n . (n-1) . (n-2) . .... (n-k+1)}{k!} \frac{\lambda^k}{n^k} \biggl(1 - \frac{\lambda}{n} \biggr)^{n} \biggl(1 - \frac{\lambda}{n} \biggr)^{-k}
\]

\[
= \frac{\lambda^k}{k!} . \frac{n . (n-1) . (n-2) . .... (n-k+1)}{n^k} . \biggl(1 - \frac{\lambda}{n} \biggr)^{n} . \biggl(1 - \frac{\lambda}{n} \biggr)^{-k}
\]

\[
\text{Here, } \frac{n . (n-1) . (n-2) . .... (n-k+1)}{n^k} \approx 1 \text{ since } n \rightarrow \infty \\
\text{And, } \biggl(1 - \frac{\lambda}{n} \biggr)^{-k} = 1 \text{ since }n \rightarrow \infty \\
\text{And, } \biggl(1 - \frac{\lambda}{n} \biggr)^{n} = e^{-\lambda} \text{ since } \biggl(1 + \frac{x}{n} \biggr)^{n} \rightarrow e^x \text{ as } n \rightarrow \infty
\]

\[
\text{So, } P[X=k] = \frac{\lambda^k}{k!} e^{-\lambda}
\]

### rpois

```{r}
rpois(n=10, 20)
```

### dpois

```{r}
success=0:30
lambda=5
pmf = dpois(x=success, lambda=lambda)
plot(success, 
     pmf,
     type='h',
     main=paste('Poisson Distribution - PMF (lambda', lambda, ')', sep=' '),
     xlab='# Success',
     ylab='Probability',
     lwd=3)
```

### ppois 

```{r}
success=0:30
lambda=5
cdf <- ppois(q=success, lambda=lambda)
plot(success, 
     cdf,
     type='h',
     main='Poisson Distribution - CDF',
     xlab='# Success',
     ylab='Probability',
     lwd=3)

```

### qpois

```{r}
quantiles <- seq(0, 1, .001)
lambda <- 5
inverse_cdf <- qpois(p=quantiles, lambda=lambda)
plot(quantiles, 
     inverse_cdf,
     type='s',
     main='Poisson Distribution - Quantile',
     xlab='p',
     ylab='Q(p)',
     lwd=1)
```

